Geometrization of characters of tori over non-Archimedean local fields

### 2013-07-18 shanghai.tex
minor fixes

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' in this
paper. Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$ \David{I think our notation is clearer if we use $k$
  rather than $\Fq$}; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,F)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob_{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob_{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

Character sheaves on $G$ form a category, denoted by $\CS(G)$. In
fact, $\CS(G)$ is a rigid symmetric monoidal category, so isomorphism
classes of character sheaves on $G$ form a group, denoted by
$\CSiso{G}$; this is the Picard group of the category of
character sheaves on $G$. Our application of character sheaves to
characters of algebraic tori over non-Archimedean local fields rests
on the Picard group of character sheaves on $G$, for certain smooth
group schemes $G$.

We now say a few words about the commutative group schemes locally of
finite type over $\Fq$ that appear in our geometrization of admissible
characters of algebraic tori over local fields. Let $K$ be a
non-Archimedean local field with residual field $\Fq$; we place no
restrictions on the characteristic of $K$ -- in particular, it may be
$0$ or not. Write $\OK$ for ring of integers of $K$ and $\pK$ for the
maximal ideal in $\OK$.  Let $T$ be an algebraic torus defined over
$K$. Let $\TT$ be a N\'eron model for $T$; recall that this is a
smooth group scheme locally of finite type over $\OK$. For each
non-negative integer $d$, let $\GN{T}_d$ be the Greenberg transform of
$\TT \otimes \OK/\pK^d$; the commutative group scheme $\GN{T}_d$
plays the r\^{o}le of $G$, above.

The main result of this paper is a canonical isomorphism
\[
\CSiso{\mathswab{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of character sheaves on
$\mathswab{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \Clifton{I'll come back to working on
  this when we're closer to being finished. I'd be just as happy if
  you took this cup from my lips, though.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasi-characters of
%$K^\times$: the monoidal category of character sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

@@

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi-2}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
\[
\begin{tikzcd}
\CS(0) \rar & \CS(\pi_0(G)) \rar & \CS(G) \rar & \CS(G^0) \rar & \CS(0)
\end{tikzcd}
\]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
\begin{equation}\label{eq:pi-1}
\begin{tikzcd}
\CSiso{0} \rar & \CSiso{\pi_0(G)} \rar & \CSiso{G} \rar & \CSiso{G^0} \rar & \CSiso{0}.
\end{tikzcd}
\end{equation}
It is easy to see that $\CSiso{0}$ is trivial. In this paper we will find $\CSiso{G^0}$ and $\CSiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated, and we will also see that \eqref{eq:pi-1} is split exact; this allows us to find $\CSiso{G}$. To prove this, we find it useful to offer an alternate perspective on a subcategory of $\CS(G)$, in the next section.

@@

In this section we describe a full subcategory of $\CS(G)$, called the category of bounded character sheaves on $G$ and denoted by $\CS_b(G)$, which will illuminate the nature of $\CS(G)$ and will also play a role in the arguments later in this paper.

>>>>

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' here. 
Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,F)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob_{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob_{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

Character sheaves on $G$ form a category, denoted by $\CS(G)$. In
fact, $\CS(G)$ is a rigid symmetric monoidal category, so isomorphism
classes of character sheaves on $G$ form a group, denoted by
$\CSiso{G}$; this is the Picard group of the category of
character sheaves on $G$. Our application of character sheaves to
characters of algebraic tori over non-Archimedean local fields rests
on the Picard group of character sheaves on $G$, for all smooth
commutative group schemes $G$, locally of finite type over $\Fq$, 
for which the geometric component group is finitely generated.

We now say a few words about the group schemes 
that appear in our geometrization of admissible
characters of algebraic tori over local fields. Let $K$ be a
non-Archimedean local field with residual field $\Fq$; we place no
restrictions on the characteristic of $K$ -- in particular, it may be
$0$ or not. Write $\OK$ for ring of integers of $K$ and $\pK$ for the
maximal ideal in $\OK$.  Let $T$ be an algebraic torus defined over
$K$. Let $\TT$ be a N\'eron model for $T$; recall that this is a
smooth group scheme locally of finite type over $\OK$. For each
positive integer $d$, let $\GN{T}_d$ be the Greenberg transform of
$\TT \otimes \OK/\pK^d$; the commutative group scheme $\GN{T}_d$
plays the r\^{o}le of $G$, above.

The main result of this paper is a canonical isomorphism
\[
\CSiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of character sheaves on
$\GN{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \Clifton{I'll come back to working on
  this when we're closer to being finished. I'd be just as happy if
  you took this cup from my lips, though.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasi-characters of
%$K^\times$: the monoidal category of character sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

@@

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi-2}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
\[
\begin{tikzcd}
\CS(0) \rar & \CS(\pi_0(G)) \rar & \CS(G) \rar & \CS(G^0) \rar & \CS(0)
\end{tikzcd}
\]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
\begin{equation}\label{eq:pi-1}
\begin{tikzcd}
\CSiso{0} \rar & \CSiso{\pi_0(G)} \rar & \CSiso{G} \rar & \CSiso{G^0} \rar & \CSiso{0}.
\end{tikzcd}
\end{equation}
It is easy to see that $\CSiso{0}$ is trivial. In this paper we will find $\CSiso{G^0}$ and $\CSiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated, and we will also see that \eqref{eq:pi-1} is split exact; this allows us to find $\CSiso{G}$.

@@

%We find it useful to offer an alternate perspective on a subcategory of $\CS(G)$.
In this section we describe a full subcategory of $\CS(G)$, called the category of bounded character sheaves on $G$ and denoted by $\CS_b(G)$, which will illuminate the nature of $\CS(G)$ and will also play a role in the arguments later in this paper.
